Verifying synchronous distributed algorithms using IVy

### 2018-08-22 paper/paper.tex
implemented comments3

Lines 2 to 5 resemble the snippet of section \textit{vii.1}. In line 2 we choose a non-determined number of nodes to fail. The \textit{assume} in line 3 then excludes all cases where a node that already failed was chosen to fail again. Line 4 adds the newly failed nodes to the set of failed nodes. Line 5 enforces $f < n$. Since this action is only called once per round we do not need an extra relation \textit{failing\_now} presented in \textit{vii.1}.

@@

\subsection{Equivalence of specification and algorithm}
One can not trivially determine, that the specification in the IVy language is logically equivalent to the actual algorithm the user wishes to verify. Whether the algorithm to be verified exists in the user's mind, in the form of pseudo-code or fully implemented in a programming language, no guarantee can be given in regards to the correctness of the translation of the algorithm into the IVy language. Some examples illustrating this difficulty will be given in the following paragraphs.

\subsubsection{Structural difference of IVy and procedural languages}
The IVy language is structurally unlike classical procedural languages. This makes it inherently challenging to achieve a one-to-one translation from a classical distributed algorithm, that was specified in a procedural way, like \textit{Floodset}. An example of this structural difference is the relatively complicated quantifier-logic in assignments that update the global network state. Take as an example the first line of \textit{step2}:

@@

As these two lines are so different, equivalence may not be obvious to a user. Proving equivalence of pseudo-code and IVy language specification would be a non-trivial proof exercise that certainly trumps manually proving \textit{Floodset}'s correctness in difficulty.

\subsubsection{Lack of modularity}
Furthermore, it has already been shown in section \textit{IV. i.1} that the lack of easy iteration over types makes it difficult to achieve modularity. The lack of which introduces additional uncertainty about correct modelling, especially in distributed systems.

\subsection{Accidental invalidation of legal state configurations with \textit{assume}}
$assume$ statements may silently invalidate state configurations. They therefore pose a threat as they
may allow false-positive IVy checks.
To give a practical example, the author of this paper originally modeled the line

\subsection{Miss-specification of system context}
As has been pointed out, it is not just the algorithm that must be correctly specified. The underlying \textit{framework} is equally important to get right. In the presented specification it is possible to remove the \textit{simulate\_failure} action and still receive a positive check from the IVy system. This is unsurprising, as we do not have any mechanism in place verifying that the system allows nodes to fail. More subtly, a user may incorrectly simulate partial message delivery of failing nodes, however all checks would still pass.

@@

\subsection{Necessity of deep understanding}
It has become apparent, that successful verification with IVy is not possible without deep knowledge about how the algorithm functions. IVy is not a ``translate and go'' solution.
Evidence of this include the complex invariant that requires knowledge about \textit{lemma 1} or the alternative encoding for $f<n$ and the decision condition $if\ rounds = f+1$.

\subsection{Algorithm failure in theory and practice}
As a user it is important to understand that IVy's goal is to verify algorithmic correctness on a high level. It does not concern itself with implementation details. However, a variety of bugs and exploits do not stem from logically unsound algorithmic foundations but implementation errors. A wonderful example is the Heartbleed exploit in OpenSSL. It would have been no help to verify the Heartbeat protocol in IVy before implementing it in C, since the error was in an implementation detail allowing a buffer underflow. The IVy user must realise, that the implementation of the algorithm must be tested through unit testing and/or bounded model checkers, despite successful verification in IVy.

\subsection{Time of specification and practicability}
Unfortunately the author missed the opportunity to scientifically record the time commitment of verifying \textit{Floodset} in IVy. He estimates the time from starting to learn IVy to successfully proving \textit{Floodset} at around 150 hours. This surprisingly large number can be attributed to a variety of factors:
\begin{itemize}
  \item Willingness to write a modular specification utilizing iteration.
  \item Attempting to count failed nodes through iteration.
  \item Having to discover an alternative encoding to $f < n$.
  \item Having to discover an alternative \textit{decision condition}.
  \item The believe that a superficial understanding of the algorithm is enough, as the proof is ``automatic'', which in turn led to a considerable amount of time investment to discover the invariants.
\end{itemize}
While drawing general conclusions about the practicability of IVy from the authors work is not possible, an attempt to quantitatively evaluate on the practicability is made in the following paragraphs. The reader is urged to read these thoughts with an increased amount of scepticism.

@@

@@

Therefore, if a user is capable of navigating the limitations with care, IVy is absolutely adequate for proving the logical soundness of an algorithm.

>>>>

Lines 2 to 5 resemble the snippet of section \textit{vii.1}. In line 2 we non-deterministically choose a number of nodes to fail. The \textit{assume} in line 3 then excludes all cases where a node that already failed was chosen to fail again. Line 4 adds the newly failed nodes to the set of failed nodes. Line 5 enforces $f < n$. Since this action is only called once per round we do not need an extra relation \textit{failing\_now} presented in \textit{vii.1}.

@@

\subsection{Specification challenges}

It has become apparent that correctly specifying an algorithm in IVy is not an easy feat. This sections summarizes and evaluates on the variety of challenges concerning correct specification that were mentioned in this paper.

\subsubsection{Equivalence of specification and algorithm}
One can not trivially determine, that the specification in the IVy language is logically equivalent to the actual algorithm the user wishes to verify. Whether the algorithm to be verified exists in the user's mind, in the form of pseudo-code or fully implemented in a programming language, no guarantee can be given in regards to the correctness of the translation of the algorithm into the IVy language. Some examples illustrating this difficulty will be given in the following paragraphs.\\

\noindent \textbf{Structural difference of IVy and procedural languages}\\
The IVy language is structurally unlike classical procedural languages. This makes it inherently challenging to achieve a one-to-one translation from a classical distributed algorithm, that was specified in a procedural way, like \textit{Floodset}. An example of this structural difference is the relatively complicated quantifier-logic in assignments that update the global network state. Take as an example the first line of \textit{step2}:

@@

As these two lines are so different, equivalence may not be obvious to a user. Proving equivalence of pseudo-code and IVy language specification would be a non-trivial proof exercise that certainly trumps manually proving \textit{Floodset}'s correctness in difficulty. \\

\noindent \textbf{Lack of modularity}\\
Furthermore, it has already been shown in section \textit{IV. i.1} that the lack of easy iteration over types makes it difficult to achieve modularity. The lack of which introduces additional uncertainty about correct modelling, especially in distributed systems.

\subsubsection{Miss-specification of system context}
As has been pointed out, it is not just the algorithm that must be correctly specified. The underlying \textit{framework} is equally important to get right. In the presented specification it is possible to remove the \textit{simulate\_failure} action and still receive a positive check from the IVy system. This is unsurprising, as we do not have any mechanism in place verifying that the system allows nodes to fail. More subtly, a user may incorrectly simulate partial message delivery of failing nodes, however all checks would still pass.

@@

\subsubsection{Necessity of deep understanding}
It has become apparent, that successful specification with IVy is not possible without deep knowledge about how the algorithm functions. IVy is not a ``translate and go'' solution.
Evidence include the alternative encoding for $f<n$ and for the decision condition $if\ rounds = f+1$.

\subsection{Verification challenges}

The following sections summarizes and evaluates on the variety of challenges concerning verification that were mentioned in this paper.

\subsubsection{Accidental invalidation of legal state configurations with \textit{assume}}
$assume$ statements may silently invalidate state configurations. They therefore pose a threat as they
may allow false-positive IVy checks.
To give a practical example, the author of this paper originally modeled the line

\subsubsection{Hard to discover inductive invariants}
Successfully discovering inductive invariants for \textit{Floodset} required a flawless understanding of the principles underlying the algorithm. Evidence of this includes the complex invariant that requires knowledge about \textit{lemma 1}. Users must be prepared to gain strong foundational understanding of the reasons why an algorithm works before attempting to find inductive invariants for it.

\subsubsection{Manual effort and practicability}
Unfortunately the author missed the opportunity to scientifically record the time commitment of verifying \textit{Floodset} in IVy. He estimates the time from starting to learn IVy to successfully proving \textit{Floodset} at around 150 hours. This surprisingly large number can be attributed to a variety of factors:
\begin{itemize}
  \item Willingness to write a modular specification utilizing iteration.
  \item Attempting to count failed nodes through iteration.
  \item Having to discover an alternative encoding to $f < n$.
  \item Having to discover an alternative \textit{decision condition}.
  \item The believe that a superficial understanding of the algorithm is enough, as the proof is ``automatic'', which in turn led to a considerable amount of time investment to discover the invariants.
\end{itemize}
While drawing general conclusions about the practicability of IVy from the authors work is not possible, an attempt to quantitatively evaluate on the practicability is made in the following paragraphs. The reader is urged to read these thoughts with an increased amount of scepticism.

@@

\subsubsection{Algorithm failure in theory and practice}
As a user it is important to understand that IVy's goal is to verify algorithmic correctness on a high level. It does not concern itself with implementation details. However, a variety of bugs and exploits do not stem from logically unsound algorithmic foundations but implementation errors. A wonderful example is the Heartbleed exploit in OpenSSL. It would have been no help to verify the Heartbeat protocol in IVy before implementing it in C, since the error was in an implementation detail allowing a buffer underflow. The IVy user must realise, that the implementation of the algorithm must be tested through unit testing and/or bounded model checkers, despite successful verification in IVy.

@@
